---
title: "Take-home Exercise 2"
author: "Sarthak Nagapurkar"
date: "May 04, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 1. Overview

This exercise aims to provide critique and improvements to visualizations created by peer in Take-home Exercise 1 using R. This will be done based on clarity and aesthetics. The original design will then be remade using data visualization design principles and best practices using `ggplot2`, its extensions, and `tidyverse` packages.

The original dataset consisted of transaction data of [REALIS](https://www.ura.gov.sg/reis/index). A complete set of the private residential property transaction data of Singapore from 1st January 2023 to 31st March 2024 was used.

## 2. Data Visualization Critique

### 2.1 Data Preparation

```{r}
pacman::p_load(ggrepel, patchwork, ggthemes, ggridges, hrbrthemes, colorspace,ggdist, tidyr, tidyverse, htmltools, viridis)
```

```{r}
realis_data_2023Q1 <- read_csv("data/ResidentialTransaction20240308160536.csv")
realis_data_2023Q1 <- mutate(realis_data_2023Q1, Quarter='2023-Q1')
realis_data_2023Q2 <- read_csv("data/ResidentialTransaction20240308160736.csv")
realis_data_2023Q2 <- mutate(realis_data_2023Q2, Quarter='2023-Q2')
realis_data_2023Q3 <- read_csv("data/ResidentialTransaction20240308161009.csv")
realis_data_2023Q3 <- mutate(realis_data_2023Q3, Quarter='2023-Q3')
realis_data_2023Q4 <- read_csv("data/ResidentialTransaction20240308161109.csv")
realis_data_2023Q4 <- mutate(realis_data_2023Q4, Quarter='2023-Q4')
realis_data_2024Q1 <- read_csv("data/ResidentialTransaction20240414220633.csv")
realis_data_2024Q1 <- mutate(realis_data_2024Q1, Quarter='2024-Q1')
df <- rbind(realis_data_2023Q1, realis_data_2023Q2, realis_data_2023Q3, realis_data_2023Q4, realis_data_2024Q1)

df <- df %>%
  mutate_if(is.character, as.factor)
```

```{r}
df$`Sale Date` = as.Date(df$`Sale Date`, format = "%d %b %Y")
```

### 2.2 Original Design

```{r}
#| fig-width: 6
#| fig-height: 6
df %>%
  filter(`Number of Units` > 1) %>%
  arrange(desc(`Number of Units`)) 
df.c <- df %>%
  filter(!grepl('ENBLOC', df$Address))
df.c <- df.c %>%
  filter(`Number of Units` <= 1)
df.c_median <- df.c %>%
  select(`Transacted Price ($)`, `Type of Area`, `Property Type`) %>%
  group_by(`Property Type`, `Type of Area`) %>%
  summarise(`Median Transaction Price ($)` = median(`Transacted Price ($)`, .groups = 'drop'))
  ggplot(df.c_median, aes(x = `Property Type`, 
             y = `Median Transaction Price ($)`, 
             color = `Type of Area`)) +  # Adding color aesthetic based on Type of Area
  geom_point() +
  theme_economist() +
  theme(axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(labels = scales::comma) +
  labs(color = "Type of Area")
```

### 2.3 Critique: Clarity and Aesthetics

#### ***2.2.1 Clarity***

-   Lack of a title makes it difficult to understand what the plot is about.
-   Dot plot is being used to represent discrete data. The differences between prices are a bit difficult to understand as dot plots are not a very good way to represent discrete data.
-   As a person who wants to understand more from the data, he will be interested in knowing the distribution of data along with the median prices as well. Just the median prices would be insufficient to derive a conclusion about the larger dataset without knowing the distribution of data.
-   Lack of subtitle makes it difficult for the person viewing the visualization to understand the message the analyst wants to convey through the visualization. In turn, it makes it difficult to understand the analyst's intention behind construction of the visualization
-   Lack of caption makes it hard to know about the source of the data.

#### ***2.2.2 Aesthetics***

-   The colour of dots for Strata matches closely with the background theme. This makes it slightly difficult to separate out.
-   The x axis labels are at 90 degrees to the usual view. This makes it uncomfortable for the reader to read the x-axis labels.
-   Y-axis gridlines make it easy to read the numbers.
-   The Type of Area has been placed at the top of the graph(in turn replacing the title). This is slightly misleading for the reader.
-   A lighter colour could have been chosen for the background. This would have made the visualization stand out more clearly.

## 3. Makeover Design

::: panel-tabset
## The plot

```{r}
#| fig-width:  11
#| fig-height: 12
#| echo: false
custom_palette <- c("orange", "steelblue")
p <- ggplot(df.c_median, aes(x = `Property Type`, 
             y = `Median Transaction Price ($)`, 
             fill = `Type of Area`)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.9) +
  theme(axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 25, face = "bold")) +
 scale_fill_manual(values = custom_palette) +
  labs(color = "Type of Area", title = "Sale Prices of Private Property types with Type of Area") + 
  scale_x_discrete(drop = FALSE)

df_terracehouse <- df.c %>% filter(`Property Type` == 'Terrace House')
p1 <- ggplot(df_terracehouse, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Terrace Houses",caption = "Source: REALIS, URA") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) +
  geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))

df_semidetached <- df.c %>% filter(`Property Type` == 'Semi-Detached House')

p2 <- ggplot(df_semidetached, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Semi-Detached Houses") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))
  
df_execcondo <- df.c %>% filter(`Property Type` == 'Executive Condominium')

p3 <- ggplot(df_execcondo, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`,fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Executive Condominium") + scale_fill_manual(values = "steelblue") +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  scale_x_discrete(drop = FALSE) +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))
  
  
df_detachedhouse <- df.c %>% filter(`Property Type` == 'Detached House')

p4 <- ggplot(df_detachedhouse, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Detached house") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))

df_condo <- df.c %>% filter(`Property Type` == 'Condominium')

p5 <- ggplot(df_condo, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Condominium") + scale_fill_manual(values = "steelblue") +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") + 
  scale_x_discrete(drop = FALSE) +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))

df_apartment <- df.c %>% filter(`Property Type` == 'Apartment')

p6 <- ggplot(df_apartment, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Apartment") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_text(size = rel(1.2)),
    axis.text.y = element_text(size = rel(0.9)),
    
    plot.background=element_rect(fill="white",colour="white"))

p7 <- p / (p6 |  p5 | p4 | p3 | p2 | p1)
p7
```

## The code

```{r}
#| fig-width:  11
#| fig-height: 12
#| eval: false
custom_palette <- c("orange", "steelblue")
p <- ggplot(df.c_median, aes(x = `Property Type`, 
             y = `Median Transaction Price ($)`, 
             fill = `Type of Area`)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.9) +
  theme(axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title = element_text(size = 25, face = "bold")) +
 scale_fill_manual(values = custom_palette) +
  labs(color = "Type of Area", title = "Sale Prices of Private Property types with Type of Area") + 
  scale_x_discrete(drop = FALSE)

df_terracehouse <- df.c %>% filter(`Property Type` == 'Terrace House')
p1 <- ggplot(df_terracehouse, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Terrace Houses",caption = "Source: REALIS, URA") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) +
  geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))

df_semidetached <- df.c %>% filter(`Property Type` == 'Semi-Detached House')

p2 <- ggplot(df_semidetached, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Semi-Detached Houses") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))
  
df_execcondo <- df.c %>% filter(`Property Type` == 'Executive Condominium')

p3 <- ggplot(df_execcondo, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`,fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Executive Condominium") + scale_fill_manual(values = "steelblue") +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  scale_x_discrete(drop = FALSE) +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))
  
  
df_detachedhouse <- df.c %>% filter(`Property Type` == 'Detached House')

p4 <- ggplot(df_detachedhouse, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Detached house") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))

df_condo <- df.c %>% filter(`Property Type` == 'Condominium')

p5 <- ggplot(df_condo, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Condominium") + scale_fill_manual(values = "steelblue") +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") + 
  scale_x_discrete(drop = FALSE) +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.background=element_rect(fill="white",colour="white"))

df_apartment <- df.c %>% filter(`Property Type` == 'Apartment')

p6 <- ggplot(df_apartment, 
             aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`)) + stat_halfeye(aes(justification = ifelse(`Type of Area` == "Land", 1.2, 0),
               side = ifelse(`Type of Area` == "Land", "left", "right")),
               adjust = 8,
               .width = 0,
               point_colour = NA, alpha = 0.9) +
  labs(x = "Apartment") + scale_fill_manual(values = custom_palette) +
  ylim(0, 15000000) + geom_boxplot(aes(x= `Type of Area`, y = `Transacted Price ($)`, fill = `Type of Area`),
               outlier.shape = NA, 
               alpha = .5, 
               width = .2, 
               colour = "black") +
  theme(
    legend.direction = "horizontal",
    legend.position="none",
    plot.title = element_text(size =12, face = "bold"),
    plot.subtitle = element_text(size = rel(0.8)),
    plot.caption = element_text(size = rel(0.8)),
    axis.text.x = element_text(size = rel(0.9)),
    axis.title.x = element_text(size = rel(0.9)),
    axis.title.y = element_text(size = rel(1.2)),
    axis.text.y = element_text(size = rel(0.9)),
    
    plot.background=element_rect(fill="white",colour="white"))

p7 <- p / (p6 |  p5 | p4 | p3 | p2 | p1)
p7

```
:::

## 4. Learning Points

This take-home exercise adds a dimension of critique and makeover of peer works which helps to embed best practices in data visualization. The key takeaways are:

-   It is very important to pay attention to detail

-   There is no best visualization as there might be some trade-offs between clarity and aesthetics. Adding horizontal (y-axis) grid lines in this chart presented me with a dilemma as it reduces aesthetics (by cluttering the plots), but adds additional clarity, assisting the audience to refer to the age group. This is particularly important for static visualization like this.

-   Learning to critique other peer’s works help to identify pain points, to understand best practices, and to awaken new ideas that were not surfaced during Take-home Exercise 1.
