---
title: "Hands-on Exercise 4: Part A"
author: "Sarthak Nagapurkar"
date: "May 03, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## **4.1 Learning Outcome**

In this hands-on exercise, we will gain hands-on experience on using:

-   ggstatsplot package to create visual graphics with rich statistical information,
-   performance package to visualise model diagnostics, and
-   parameters package to visualise model parameters

## **4.2 Visual Statistical Analysis with ggstatsplot**

[**ggstatsplot**](https://indrajeetpatil.github.io/ggstatsplot/)  is an extension of [**ggplot2**](https://ggplot2.tidyverse.org/) package for creating graphics with details from statistical tests included in the information-rich plots themselves.

## **4.3 Getting Started**

### **10.3.1 Installing and launching R packages**

In this exercise, **ggstatsplot** and **tidyverse** will be used.

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

### **4.3.2 Importing data**

```{r}
exam <- read_csv("data/Exam_data.csv")
```

### **4.3.3 One-sample test: *gghistostats()* method**

```{r}
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English Scores"
)
```

### **4.3.4 Unpacking the Bayes Factor**

-   A Bayes factor is the ratio of the likelihood of one particular hypothesis to the likelihood of another. It can be interpreted as a measure of the strength of evidence in favor of one theory among two competing theories.
-   That’s because the Bayes factor gives us a way to evaluate the data in favor of a null hypothesis, and to use external information to do so. It tells us what the weight of the evidence is in favor of a given hypothesis.
-   When we are comparing two hypotheses, H1 (the alternate hypothesis) and H0 (the null hypothesis), the Bayes Factor is often written as B10. It can be defined mathematically as
-   The [**Schwarz criterion**](https://www.statisticshowto.com/bayesian-information-criterion/) is one of the easiest ways to calculate rough approximation of the Bayes Factor.

### **4.3.5 How to interpret Bayes Factor**

A **Bayes Factor** can be any positive number. One of the most common interpretations is this one—first proposed by Harold Jeffereys (1961) and slightly modified by [Lee and Wagenmakers](https://www-tandfonline-com.libproxy.smu.edu.sg/doi/pdf/10.1080/00031305.1999.10474443?needAccess=true) in 2013:

### **4.3.6 Two-sample mean test: *ggbetweenstats()***

In the code chunk below, [*ggbetweenstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbetweenstats.html) is used to build a visual for two-sample mean test of Maths scores by gender.

```{r}
ggbetweenstats(
  data = exam,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)
```

Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

### **4.3.7 Oneway ANOVA Test: *ggbetweenstats()* method**

In the code chunk below, [*ggbetweenstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbetweenstats.html) is used to build a visual for One-way ANOVA test on English score by race.

```{r}
ggbetweenstats(
  data = exam,
  x = RACE,
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

-   “ns” → only non-significant
-   “s” → only significant
-   “all” → everything

### **4.3.8 Significant Test of Correlation: *ggscatterstats()***

In the code chunk below, [*ggscatterstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggscatterstats.html) is used to build a visual for Significant Test of Correlation between Maths scores and English scores.

```{r}
ggscatterstats(
  data = exam,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE
)
```

### **4.3.9 Significant Test of Association (Depedence) : *ggbarstats()* methods**

In the code chunk below, the Maths scores is binned into a 4-class variable by using [*cut()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cut).

```{r}
exam1 <- exam %>%
  mutate(MATHS_bins = 
           cut(MATHS,
               breaks= c(0,60,75,85,100)))

```

In this code chunk below [*ggbarstats()*](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbarstats.html) is used to build a visual for Significant Test of Association

```{r}
ggbarstats(exam1,
           x = MATHS_bins,
           y = GENDER)
```
