---
title: "Hands-on Exercise 1"
author: "Sarthak Nagapurkar"
---

# Getting Started

## Installing and loading R packages

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("C:/Users/DELL/Downloads/Hands-on_Ex1/data/Exam_data.csv")
```

```{r}
ggplot(data=exam_data)
```

The aesthetic mappings take attributes of the data and and use them to influence visual characteristics, such as position, colour, size, shape, or transparency. Each visual characteristic can thus encode an aspect of the data and be used to convey information.

All aesthetics of a plot are specified in the [`aes()`](https://ggplot2.tidyverse.org/reference/aes.html) function call

Code chunk below adds the aesthetic element into the plot.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

# Plotting a Bar chart

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

# Plotting a Dot plot

In a dot plot, the width of a dot corresponds to the bin width (or maximum width, depending on the binning algorithm), and dots are stacked, with each dot representing one observation.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

The `scale_y_continuous()` is used to turn off the y-axis, and *binwidth argument* is used to change the binwidth to 2.5.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

# Plotting a histogram

Plot a histogram for Maths scores

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()       
```

Modify the geometric object by changing the geom() object

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

Modify the geometric object by changing the aes() object

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

# Plotting the kernel estimate density plot

The code below plots the distribution of Maths scores in a kernel density estimate plot.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
```

The code chunk below plots two kernel density lines by using *colour* or *fill* arguments of *aes()*

```{r}
ggplot(data=exam_data,
       aes(x = MATHS,colour=GENDER)) +
  geom_density()
```

# Plotting a Boxplot

```{r}
ggplot(data=exam_data,
       aes(y= MATHS,x = GENDER)) +
  geom_boxplot()
```

The code chunk below plots the distribution of Maths scores by gender in notched plot instead of boxplot.

```{r}
ggplot(data=exam_data,
       aes(y = MATHS,x = GENDER)) +
  geom_boxplot(notch=TRUE)
```

# Plotting a Violin Plot

The code below plot the distribution of Maths score by gender in violin plot.

```{r}
ggplot(data=exam_data,
       aes(y=MATHS, x=GENDER)) + 
  geom_violin()
```

# Plotting a Scatter Plot

The code chunk below plots a scatterplot showing the Maths and English grades of pupils.

```{r}
ggplot(data=exam_data,
       aes(x=MATHS, y=ENGLISH)) +
  geom_point()
```

## Combining Boxplots and Scatter plots

The code chunk below plots the data points on the boxplots by using both `geom_boxplot()` and `geom_point()`.

```{r}
ggplot(data=exam_data,
       aes(y=MATHS, x=GENDER)) +
  geom_point(position="jitter", size=0.5) + 
  geom_boxplot()
```

# Working with stat

## **Working with stat - the *stat_summary()* method**

The code chunk below adds mean values by using [`stat_summary()`](https://ggplot2.tidyverse.org/reference/stat_summary.html) function and overriding the default geom.

```{r}
ggplot(data=exam_data,
       aes(y=MATHS, x=GENDER)) +
  geom_boxplot() +
  stat_summary(geom= "point",
               fun= "mean",
               colour= "red",
               size= 4)
```

##  **Working with stat - the `geom()` method**

The code chunk below adding mean values by using `geom_()` function and overriding the default stat.

```{r}
ggplot(data=exam_data,
       aes(y=MATHS, x=GENDER)) +
  geom_boxplot() + 
  geom_point(stat="summary",
             fun= "mean",
             colour= "red",
             size= 4)
```

## Adding a best fit curve on a scatter plot

The scatterplot below shows the relationship of Maths and English grades of pupils. The interpretability of this graph can be improved by adding a best fit curve.In the code chunk below, [`geom_smooth()`](https://ggplot2.tidyverse.org/reference/geom_smooth.html) is used to plot a best fit curve on the scatterplot.

```{r}
ggplot(data=exam_data,
       aes(y= ENGLISH, x= MATHS)) +
  geom_point() +
  geom_smooth(linewidth=0.5)
```

The default smoothing method is loess. The default smoothing method can be overridden as shown below.

```{r}
ggplot(data= exam_data,
       aes(y= ENGLISH, x= MATHS)) +
  geom_point() +
  geom_smooth(method= lm,
              linewidth= 0.5)
```

# Working with facets

## **Working with `facet_wrap()`**

[`facet_wrap`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than facet_grid because most displays are roughly rectangular. The code chunk below plots a trellis plot using `facet-wrap()`.

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 25) +
  facet_wrap(~ CLASS)
```

## Working with facet_grid()

[`facet_grid()`](https://ggplot2.tidyverse.org/reference/facet_grid.html) forms a matrix of panels defined by row and column facetting variables. It is most useful when you have two discrete variables, and all combinations of the variables exist in the data. The code chunk below plots a trellis plot using `facet_grid()`.

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) + 
  geom_histogram(bins = 25) + 
  facet_grid(~ CLASS)
```

# Working with Coordinate

By the default, the bar chart of ggplot2 is in vertical form.

```{r}
ggplot(data = exam_data,
       aes(x=RACE)) +
  geom_bar()
```

The code chunk below flips the horizontal bar chart into vertical bar chart by using `coord_flip()`.

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar() +
  coord_flip()
```

## Changing the x and y-axis range

The scatter plot below is slightly misleading because the y-axis and x-axis range are not equal.

```{r}
ggplot(data = exam_data,
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, linewidth=0.5)
```

The code chunk below fixed both the y-axis and x-axis range from 0-100.

```{r}
ggplot(data = exam_data,
       aes(x = MATHS, y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, linewidth=0.5) +
  coord_cartesian(xlim = c(0, 100),
                  ylim = c(0, 100))
```
