---
title: "Hands-on Exercise 1"
author: "Sarthak Nagapurkar"
---

# Getting Started

## Installing and loading R packages

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("C:/Users/DELL/Downloads/Hands-on_Ex1/data/Exam_data.csv")
```

```{r}
ggplot(data=exam_data)
```

The aesthetic mappings take attributes of the data and and use them to influence visual characteristics, such as position, colour, size, shape, or transparency. Each visual characteristic can thus encode an aspect of the data and be used to convey information.

All aesthetics of a plot are specified in the [`aes()`](https://ggplot2.tidyverse.org/reference/aes.html) function call

Code chunk below adds the aesthetic element into the plot.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

# Plotting a Bar chart

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

# Plotting a Dot plot

In a dot plot, the width of a dot corresponds to the bin width (or maximum width, depending on the binning algorithm), and dots are stacked, with each dot representing one observation.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

The `scale_y_continuous()` is used to turn off the y-axis, and *binwidth argument* is used to change the binwidth to 2.5.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

# Plotting a histogram

Plot a histogram for Maths scores

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()       
```

Modify the geometric object by changing the geom() object

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

Modify the geometric object by changing the aes() object

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

# Plotting the kernel estimate density plot

The code below plots the distribution of Maths scores in a kernel density estimate plot.

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
```

The code chunk below plots two kernel density lines by using *colour* or *fill* arguments of *aes()*

```{r}
ggplot(data=exam_data,
       aes(x = MATHS,colour=GENDER)) +
  geom_density()
```

# Plotting a Boxplot

```{r}
ggplot(data=exam_data,
       aes(y= MATHS,x = GENDER)) +
  geom_boxplot()
```

The code chunk below plots the distribution of Maths scores by gender in notched plot instead of boxplot.

```{r}
ggplot(data=exam_data,
       aes(y = MATHS,x = GENDER)) +
  geom_boxplot(notch=TRUE)
```

# Plotting a Violin Plot

The code below plot the distribution of Maths score by gender in violin plot.

```{r}
ggplot(data=exam_data,
       aes(y=MATHS, x=GENDER)) + 
  geom_violin()
```

# Plotting a Scatter Plot

The code chunk below plots a scatterplot showing the Maths and English grades of pupils.

```{r}
ggplot(data=exam_data,
       aes(x=MATHS, y=ENGLISH)) +
  geom_point()
```
