---
title: "Hands-on Exercise 2"
author: "Sarthak Nagapurkar"
date: "April 17, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting Started

## **Installing and loading the required libraries**

```{r}
pacman::p_load(ggrepel, patchwork, ggthemes, hrbrthemes, tidyverse)
```

## Importing the data

The code chunk below imports *exam_data.csv* into R environment by using [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/) package. **readr** is one of the tidyverse package.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Beyond ggplot2 Annotation: ggrepel

One of the challenge in plotting statistical graph is annotation, especially with large number of data points.

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label(aes(label = ID),
             hjust = 0.5,
             vjust = -0.5) + 
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) + 
  ggtitle("English scores versus Maths scores for Primary 3")
```

## The code

```{r}
#| eval: false
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label(aes(label = ID),
             hjust = 0.5,
             vjust = -0.5) + 
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) + 
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

## Working with ggrepel

We simply replace `geom_text()` by [`geom_text_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel.html) and `geom_label()` by [`geom_label_repel`](https://ggrepel.slowkow.com/reference/geom_text_repel.html)

::: panel-tabset
## The plot

```{r}
#| echo: false
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label_repel(aes(label = ID),
             fontface="bold") + 
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) + 
  ggtitle("English scores versus Maths scores for Primary 3")
```

## The code

```{r}
#| eval: false
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm,
              size=0.5) +
  geom_label_repel(aes(label = ID),
             fontface="bold") + 
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) + 
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::
